A prototype software framework for transferable computational health economic models and its early application in youth mental health

### 2024-02-21 Markdown/Manuscript/Child_RMDs/MS_Aims.Rmd
minor tidies

(i)	a prototype software framework we have developed to support the development of modular and open source CHEMs; and
<!-- propose a set of ethical responsibilities for CHEM project teams and criteria for assessing ethical CHEM implementations; -->
<!-- (ii)  -->
(ii)	how we have applied that software framework to develop and share an initial set of modules for a economic model of youth mental health.

>>>>

(i)	our motivation for developing a modular and open source CHEM in youth mental health;

(ii) a prototype software framework we have developed to support the development of transferable CHEMs; and
<!-- propose a set of ethical responsibilities for CHEM project teams and criteria for assessing ethical CHEM implementations; -->
<!-- (ii)  -->
(ii)	how we have applied the prototype software framework to develop and share an initial set of modules for our youth mental health CHEM.

### 2024-02-21 Markdown/Manuscript/Child_RMDs/MS_Imp_Motivation.Rmd
minor tidies

Potential future directions are to supplement readyforwhatsnext with projects in two additional Wagstaff and Culyer domains of: (i) public health (to model the impact of selected fiscal policy and regulation options on young people's mental health); and (ii) human resources (to model the supply and behaviours of the youth mental health workforce). Our ultimate aim is for readyforwhatsnext to flexibly combine its constituent parts in analyses that help answer questions in two additional Wagstaff and Culyer domains:

@@

## User requirements
Although we are principally interested in using readyforwhatsnext to answer policy questions relating to the mental health of young people in Australia, we want to facilitate model transfer to other jurisdictions. We believe that a good starting point for facilitating model transferable is to develop a CHEM that is relatively easy to understand, use and adapt. For that reason, the initial user-requirements we specified for readyforwhatsnext are to meet six explicit criteria for transparency, reusability and updatability. We have describe these criteria elsewhere [@ethicsCHEMS2024], but briefly they are:

@@

T4) Licenses restrict only unethical reuse; <!-- rephrase Liceses permit derivative works-->

@@

T6) New releases are systematically retested, with deprecation of outdated artefacts.

>>>>

Potential future directions are to incorporate projects in two additional Wagstaff and Culyer domains of: (i) public health (to model the impact of selected fiscal policy and regulation options on young people's mental health); and (ii) human resources (to model the supply and behaviours of the youth mental health workforce). Our ultimate aim is for readyforwhatsnext to flexibly combine its constituent parts in analyses that help answer questions in two additional Wagstaff and Culyer domains:

@@

## Transferability
Although we are principally interested in using readyforwhatsnext to answer policy questions relating to the mental health of young people in Australia, we want to facilitate model transfer to other jurisdictions. We believe that a good starting point for facilitating model transferable is to develop a CHEM that is relatively easy to understand, use and adapt. For that reason, the transferability requirements we specified for readyforwhatsnext are to meet six explicit criteria for transparency, reusability and updatability. We have describe these criteria and theire relationship to ethical modelling practice elsewhere [@ethicsCHEMS2024], but briefly they are:

@@

T4) Licenses restrict permit derivative works; <!-- rephrase Liceses permit derivative works-->

@@

T6) Systematic retesting and deprecation.

### 2024-02-21 Markdown/Manuscript/Child_RMDs/MS_Intro_Short.Rmd
minor tidies

<!-- Imported_from_Ethics-->Key concerns for health economists when considering whether to reuse a model are generalisability (application without adaptation) and transferability (selective reuse and/or modification of model components) [@RN39]. <!-- End_Imported_from_Ethics -->A potential strategy for achieving both these goals is to make computational health economic models (CHEMs) both modular and open-source. A modular model is constructed from multiple reusable and replaceable sub-models (modules) [@pan2021modular]. Modules can share inputs and outputs with other modules or can be run as independent models [@barros2023empowering]. Advantages of modular models include feasibility (large projects are broken into smaller tasks, with each component independently developed and tested) and flexibility (making it easier to selectively replace or update specific parts of a model and to scale up or down the level of granularity) [@pan2021modular]. Modular approaches are currently being used to facilitate the development of complex computational models in disciplines such as biology [@pan2021modular], ecology [@barros2023empowering] and neuroscience [@frazier2022robust]. In health economics the related and enabling concept of reference models has been recommended [@Afzali2013], but peer reviewed studies describing modular health economic models remain relatively rare, though examples exist in infectious disease [@Trauer2017] and cardiology [@Urach2013].

>>>>

<!-- Imported_from_Ethics-->Key concerns for health economists when considering whether to reuse a model are generalisability (application without adaptation) and transferability (selective reuse and/or modification of model components) [@RN39]. Health economic models are likely to require at least some modification to be validly applied to a new jurisdiction, time period or decision problem. How difficult it is to undertake such a model transfer depends in part on whether the Computational Health Economic Model (CHEM) - the software files that implement the model - can be readily understood, applied and adapted.

<!-- End_Imported_from_Ethics -->A potential strategy for promoting the transferability of health economic models is to develop CHEMs that are both modular and open-source. A modular model is constructed from multiple reusable and replaceable sub-models (modules) [@pan2021modular]. Modules can share inputs and outputs with other modules or can be run as independent models [@barros2023empowering]. Advantages of modular models include feasibility (large projects are broken into smaller tasks, with each component independently developed and tested) and flexibility (making it easier to selectively replace or update specific parts of a model and to scale up or down the level of granularity) [@pan2021modular]. Modular approaches are currently being used to facilitate the development of complex computational models in disciplines such as biology [@pan2021modular], ecology [@barros2023empowering] and neuroscience [@frazier2022robust]. In health economics the related and enabling concept of reference models has been recommended [@Afzali2013], but peer reviewed studies describing modular health economic models remain relatively rare, though examples exist in infectious disease [@Trauer2017] and cardiology [@Urach2013].

### 2024-02-21 Markdown/Manuscript/Parent_PDF/Springer.Rmd
minor tidies

title: A prototype software framework for transferable computational health economic models
titlerunning: Framework for transferable computational models

>>>>

title: A prototype software framework for transferable computational health economic models and its early application in youth mental health
titlerunning: Framework for transferable computational models
